### 2018-08-30 smt/ivanova.tex
MI: created folder with tex sources of revised version of SMT

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[natbib,smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
\bibliographystyle{aps-nameyear}
%
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{subcaption}
\usepackage{hyperref}
%\usepackage{pdflscape} % Landscape pages
\usepackage{color}	% Different font colors
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage[]{algorithm2e}
\usepackage{color}
\usepackage{multirow} % Merging rows in tables
\usepackage{multicol} % Merging columns in tables
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{mathrsfs} % Fancy math font for model labels
\usepackage{tikz}

\newcommand{\indentitem}{\setlength\itemindent{18pt}}
\renewcommand{\labelitemi}{$\bullet$}
\captionsetup{compatibility=false}
\newtheorem{prop}{Proposition}
\newtheorem{obs}{Observation}
\newtheorem{cor}{Corollary}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
\SetCommentSty{mycommfont}
\SetKwRepeat{Do}{do}{while}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{} 
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
% original amsmath definition
% \subequations:
% \long macro:->\refstepcounter {equation}\protected@edef \theparentequation {\theequation }\setcounter {parentequation}{\value {equation}}\setcounter {equation}{0}\def \theequation {\theparentequation \alph {equation}}\ignorespaces

% saved by hyperref in
% > \HyOrg@subequations=\long macro:

% hyperref-patched \subequations: (\endsubequations not modified)
% > \subequations=macro:
% ->\stepcounter {equation}\protected@edef \theHparentequation {\@ifundefined {th
% eHequation}\theequation \theHequation }\addtocounter {equation}{-1}\HyOrg@subeq
% uations \def \theHequation {\theHparentequation \alph {equation}}\ignorespaces 
% .

% extending the environment
%   1. with optional parameter: expected values resume or intermezzo or none.
%   2. while keeping the hyperref customization.

\makeatletter
\def\user@resume{resume}
\def\user@intermezzo{intermezzo}
%
\newcounter{previousequation}
\newcounter{lastsubequation}
\newcounter{savedparentequation}
\setcounter{savedparentequation}{1}
% 
% The following code causes that we are able to continue with numbering of a model in a next \align .
% For example:
%    a < b    (1a)
%    a < c    (1b)
% some text comes here and the numbering will continue:
%    a < d    (1c)
%
\renewenvironment{subequations}[1][]{%
      \def\user@decides{#1}%
      \setcounter{previousequation}{\value{equation}}%
      \ifx\user@decides\user@resume 
           \setcounter{equation}{\value{savedparentequation}}%
      \else  
      \ifx\user@decides\user@intermezzo
           \refstepcounter{equation}%
      \else
           \setcounter{lastsubequation}{0}%
           \refstepcounter{equation}%
      \fi\fi
      \protected@edef\theHparentequation{%
          \@ifundefined {theHequation}\theequation \theHequation}%
      \protected@edef\theparentequation{\theequation}%
      \setcounter{parentequation}{\value{equation}}%
      \ifx\user@decides\user@resume 
           \setcounter{equation}{\value{lastsubequation}}%
         \else
           \setcounter{equation}{0}%
      \fi
      \def\theequation  {\theparentequation  \alph{equation}}%
      \def\theHequation {\theHparentequation \alph{equation}}%
      \ignorespaces
}{%
%  \arabic{equation};\arabic{savedparentequation};\arabic{lastsubequation}
  \ifx\user@decides\user@resume
       \setcounter{lastsubequation}{\value{equation}}%
       \setcounter{equation}{\value{previousequation}}%
  \else
  \ifx\user@decides\user@intermezzo
       \setcounter{equation}{\value{parentequation}}%
  \else
       \setcounter{lastsubequation}{\value{equation}}%
       \setcounter{savedparentequation}{\value{parentequation}}%
       \setcounter{equation}{\value{parentequation}}%
  \fi\fi
%  \arabic{equation};\arabic{savedparentequation};\arabic{lastsubequation}
  \ignorespacesafterend
}
\makeatother

\begin{document}

\title{Integer Programming Formulations for the Shared Multicast Tree Problem
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
% \subtitle{Tightening the LP bounds}
\titlerunning{IP formulations for SMT}        % if too long for running head

\author{Marika Ivanova        \and
         Dag Haugland
}
%\authorrunning{Short form of author list} % if too long for running head

\institute{M. Ivanova \at
              Department of Informatics, University of Bergen, Norway \\
              Tel.: +47 942 77 070\\
              \email{Marika.Ivanova@uib.no}         %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           D. Haugland \at
              Department of Informatics, University of Bergen, Norway \\
              Tel.: +47 55584033\\
              \email{Dag.Haugland@uib.no}         %  \\
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor
\maketitle

\begin{abstract}
We study the Shared Multicast Tree (SMT) problem in wireless networks.
To support a multicast session between a set of network nodes, SMT aims to establish a wireless 
connection between them, such that the total energy consumption is minimized.
All destinations of the multicast message must be connected, while non-destinations are optional nodes that can be used to relay messages. 
In the well-studied minimum energy multicast (MEM)
problem, the optimal tree depends on the source of the message.
In contrast, SMT asks for a source-independent tree, which substantially reduces the processing overhead at each node.
This means that SMT can be regarded as a Steiner tree problem.
The objective function reflecting power consumption, however, distinguishes SMT clearly from the traditional minimum Steiner tree problem.

In the present work, we develop two 
integer programming formulations for SMT.
Both models are subsequently extended and strengthened by redundant valid inequalities.
Further strengthening is achieved by extending the set of variables and corresponding constraints.
Theorems on relations between the LP bounds corresponding to the models are stated and proved.
As the number of variables in the strongest formulations is a polynomial of degree four in the number of network nodes,
the models are impractical for computing lower bounds in instances beyond a fairly small size.
To make them applicable to larger instances, a constraint generation scheme is developed.
Results from computational experiments with the models demonstrate good promise of the approaches taken.
\keywords{Wireless communication, multicast tree, Steiner tree, LP bound, valid inequalities}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\input{sec1-intro.tex}
\input{sec2-sbtsmt.tex}
\input{sec3-milp.tex}
\input{sec4-relations.tex}
\input{sec5-strong.tex}
\input{sec6-exp.tex}
\input{sec7-conclusion.tex}

%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}

% For one-column wide figures use
%\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%
% For two-column wide figures use
%\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
% For tables use
%\begin{table}
% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}

%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
% \bibitem[Dian{\'e} and Plesn{\'i}k(1993)]{diane93ipf}
% Dian{\'e} M, Plesn{\'i}k J (1993)
% An integer programming formulation of the Steiner problem in graphs.
% Zeitschrift f{\"u}r Operations Research
% 37(1):107--111

\bibitem[Althaus et al.(2003)]{althaus03}
Althaus E, C\u{a}linescu G, M\u{a}ndoiu II, Prasad S, Tchervenski N, Zelikovsky A (2003)
Power efficient range assignment in ad-hoc wireless networks.
In: Proceedings IEEE WCNC’03, pp 1889--1894

\bibitem[Altinkemer et al.(2005)]{altinkemer05}
Altinkemer K, Salman FS, Bellur P (2005)
Solving the minimum energy broadcasting problem in ad hoc wireless networks by integer programming.
In: Proceedings INOC 2005, pp B2.635--B2.642

\bibitem[Bauer et al.(2008)]{bauer08}
Bauer J, Haugland D, Yuan D (2008)
Analysis and computational study of several integer programming formulations for minimum-energy multicasting in wireless ad hoc networks.
Networks 52(2):57–68

\bibitem[\v{C}agalj et al.(2002)]{cagalj02}
\v{C}agalj M, Hubaux J-P, Enz C (2002)
Minimum-energy broadcast in all-wireless networks: NP-completeness and distribution issues.
In: Proceedings ACM MOBICOM 2002 pp 172--182

\bibitem[Clementi et al.(1999)]{clementi99}
Clementi A, Penna P, Silvestri R (1999)
Hardness results for the power range assignment problem in packet radio networks.
In: Hochbaum D et al. (eds) RANDOM-APPROX’99,
LNCS 1671:197--208

\bibitem[Das et al.(2003)]{das03}
Das AK, Marks RJ, El-Sharkawi M, Arabshani P, Gray A (2003)
Minimum power broadcast trees for wireless networks: integer programming formulations.
In: Proceedings IEEE INFOCOM 2003 Vol.\ 2, pp 1001--1010

\bibitem[Das et al.(2005)]{das05}
Das AK, Marks RJ, El-Sharkawi E, Arabshahi P, Gray A (2005)
Optimization methods for minimum power bidirectional topology construction in wireless networks with sectored antennas.
In: Proceedings IEEE GLOBECOM ’05, pp 3962--3968

\bibitem[Goemans and Myung(1993)]{goemans93catalog}
Goemans MX, Myung Y-S (1993)
A catalog of Steiner tree formulations.
Networks
23(1):19--28

\bibitem[Haugland and Yuan(2011)]{Haugland11Compact}
Haugland D, Yuan D (2011)
Compact integer programming models for power-optimal trees in ad hoc wireless networks.
In: Kennington J, Olinick E, Rajan D (eds) Wireless network design - optimization models and solution procedure.
International series in operations research \& management science.
Springer, New York, pp 219--246

\bibitem[Hsiao et al.(2013)]{hsiao13}
Hsiao P-C, Chiang T-C, Fu L-C (2013)
Static and dynamic minimum energy broadcast problem in wireless ad-hoc networks: A PSO-based approach and analysis
Applied Soft Computing
13:786--4801

\bibitem[Ivanova(2016)]{ivanova16isco}
Ivanova M (2016)
Shared Multicast Trees in Ad Hoc Wireless Networks.
In: Cerulli R, Fujishige S, Mahjoub A (eds) Combinatorial Optimization ISCO 2016,
LNCS 9849:273--284

\bibitem[Kirousis et al.(1997)]{kirousis97}
Kirousis LM, Kranakis E, Krizanc D, Pelc A (1997)
Power consumption in packet radio networks.
In: Reischuk M (ed): STACS'97 Proceedings,
LNCS 1200:363--374

\bibitem[Leggieri et al.(2008)]{leggieri08}
Leggieri V, Nobili P, Triki C (2008)
Minimum power multicasting problem in wireless networks.
Math Meth Oper Res 68:295–311

\bibitem[Montemanni and Gambardella(2004)]{montemanni04}
Montemanni R, Gambardella LM (2004)
Exact algorithms for the minimum power symmetric connectivity problem in wireless networks.
Comput Oper Res 31(10):1667--1680

\bibitem[Montemanni and Leggieri(2011)]{montemanni11}
Montemanni R, Leggieri V (2011)
A branch and price algorithm for the minimum power multicasting problem in wireless sensor networks.
Math Meth Oper Res
74:327--342

\bibitem[Papadimitriou and Georgiadis(2006)]{Papadimitriou06SBT}
Papadimitriou I, Georgiadis L (2006)
Minimum-energy Broadcasting in Multi-hop Wireless Networks Using a Single Broadcast Tree.
Mobile Netw Appl
11(3):361--375

\bibitem[Polzin and Daneshmand(2001)]{Polzin}
Polzin T, Daneshmand SV (2001)
A comparison of Steiner tree relaxations.
Discrete Appl Math
112(1--3):241--261

\bibitem[Wieselthier et al.(2000)]{Wieseltier00onthe}
Wieselthier JE, Nguyen GD, Ephremides A (2000)
On the Construction of Energy-Efficient Broadcast and Multicast Trees in Wireless Networks,
In: Proceedings IEEE INFOCOM 2000 Vol.\ 2, pp 585--594

\bibitem[Yuan(2005)]{yuan05}
Yuan D (2005)
An integer programming approach for the minimum-energy broadcast problem in wireless networks.
In: Proceedings INOC 2005, pp B2.643–B2.650

\bibitem[Yuan et al.(2008)]{yuan08}
Yuan D, Bauer J, Haugland D (2008)
Minimum-energy broadcast and multicast in wireless networks: An
integer programming approach and improved heuristic algorithms.
Ad Hoc Netw
6(5):696–717

\bibitem[Yuan and Haugland(2012)]{Haugland12Dual}
Yuan D, Haugland D (2012)
Dual Decomposition for Computational Optimization of Minimum-Power Shared Broadcast Tree in Wireless Networks.
IEEE T Mobile Comput
12(11):2008--2019

%\bibitem{RefJ}
% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}
% end of file template.tex