### 2018-08-30 smt/sec2-sbtsmt.tex

\section{The Minimum Shared Multicast Tree Problem}
\label{sec:SBT}

\subsection{Assumptions and notation} \label{sec:notation}

A WANET is modeled by a complete graph $G=(V,E)$, where the set $V$ of nodes represents the set of wireless devices and the set of edges $E=\{\{i,j\}:i,j\in V, i\neq j\}$ corresponds to the potential links between them.
The set $A=\{(i,j):i,j\in V,\{i,j\}\in E\}$ consists of all arcs that can be derived by directing the edges in $E$.
Nodes in the set $D\subseteq V$ are the destination nodes, also referred to as \emph{source} nodes.
For an arbitrary $i\in V$, sets $V\setminus \{i\}$ and $D\setminus\{i\}$ are denoted $V_i$ and $D_i$, respectively.

Sending a message directly from (to) node $i$ to (from) node $j$ requires a certain \emph{power} $p_{ij}=p_{ji}$.
Whenever appropriate, $p_{ij}$ is referred to as the \emph{cost} of $\{i,j\}$,
and edges with higher cost than others are said to be \emph{more expensive}.
When the nodes are embedded in the plane, $p_{ij}$ is often assumed \citep{Papadimitriou06SBT,Haugland12Dual} to be
proportional to the Euclidean distance between $i$ and $j$ raised to a power between 2 and 4.
For simplicity, we assume for all $i\in V$ that $p_{ij}\neq p_{ik}$ when $j\neq k$, and define
$W_{ij}=\left\{k\in V_i: p_{ik}\geq p_{ij}\right\}$
as the set of nodes requiring no less power at $i$ than $j$ does.
The notation $G'\subseteq G$ means that $G'=(V',E')$ is a subgraph of $G$, i.e., $E'\subseteq E$ consists of the edges in $E$ with both end nodes in $V'\subseteq V$.

Let $y \in \{0,1\}^E$ and $X\in\{0,1\}^{A\times D}$ be binary vectors with components corresponding to
edges and pairs of arcs and destination nodes, respectively.
The vector $X^s$ then consists of all components of $X$ corresponding to a pair of which $s$ is the destination.
The subgraph of $G$ \emph{induced by} $y$ is defined as $G_y=(V,E_y)$, where $E_y=\left\{\{i,j\}\in E: y_{ij}=1\right\}$.
Directed graphs induced by binary vectors over $A$ are defined analogously.

Several integer programming models, all of which exclusively have binary variables, will be considered in this text.
If $\mathcal{M}$ is such a model, $\text{LP}(\mathcal{M})$ denotes the corresponding continuous relaxation
where the domain of each variable is replaced by the interval $[0,1]$.
Further, $z\left(\mathcal{M}\right)$ and $z\left(\text{LP}(\mathcal{M})\right)$ denote the optimal objective function values of the respective models.

\subsection{Problem definition} \label{sec:probdef}

To illustrate the problem under study, consider the tree depicted in Fig. \ref{fig:objexp}.
Six destination nodes ($D=\{i_1,i_2,a,b,c,d\}$) are spanned by use of three Steiner nodes ($V\setminus D = \{e,i,i_3\}$).
Consider node $i$ and its three neighbour nodes $i_1$, $i_2$ and $i_3$, ordered by decreasing power requirement at $i$.
That is, $p_{ii_1}>p_{ii_2}>p_{ii_3}$.
If the transmitting node is outside the shaded area ($i_2$, $c$, or $d$), then node $i$ relays the signal to two of its neighbors, \emph{including $i_1$}.
Exploiting the wireless advantage, node $i$ reaches both neighbors if it is assigned power level $p_{ii_1}$.
Otherwise, if the signal source is either $i_1$, $a$, or $b$, node $i$ passes the signal on to neighbors $i_2$ and $i_3$.
This is accomplished by power $p_{ii_2}$.

\begin{figure}[h!]
\centering
\begin{tikzpicture}[scale=.6]
 \draw[rounded corners=5mm, fill=gray!20, draw=white] (-.7,-.2) -- (3.5,0.7) -- (0.2,4.2) -- cycle;
 \begin{scope}[every node/.style={circle,draw=black,fill=gray!60,minimum size=1.5em, inner sep=2, scale=.8}]
  \node (a) at (0,0.5) {$a$};
  \node (b) at (0.5,3) {$b$};
  \node (c) at (9,2.5) {$c$};
  \node (d) at (9.5,0.9) {$d$};
  \node[scale=.85] (i1) at (2,1) {$i_1$};
  \node[scale=.85] (i2) at (4,3.1) {$i_2$};
  \node[scale=.85,fill=none] (i3) at (5.8,1.2) {$i_3$};
  \node[fill=none] (i) at (4.5,1.5) {$i$};
  \node[fill=none] (e) at (8,1.5) {$e$};
 \end{scope}
 \draw (a) -- (i1) -- (b);
 \draw (i1) -- (i) -- (i3) -- (e) -- (c);
 \draw (e) -- (d);
 \draw (i) -- (i2);
\end{tikzpicture}
\caption{A multicast tree illustrating the cost contribution from node $i$ to the objective function.
	 Destinations and Steiner nodes are denoted by filled and empty circles, respectively.}
\label{fig:objexp}
\end{figure}

In more general terms, let $T$ be a multicast tree, and let $i\in V_T$ be a node with at least two neighbors in $T$.
For multicasting a message, the power to be assigned to node $i$ is either $p_{ii_1}$ or $p_{ii_2}$, where $i_1$ and $i_2$ are neighbors of $i$ such that
$p_{ii_1}>p_{ii_2}>p_{ij}$ for all other neighbors $j\in V_T\setminus\{i_1,i_2\}$.
If the path from the message source to $i$ intersects $i_1$, then power $p_{ii_2}$ is sufficient.
Otherwise, the power $p_{ii_1}$ is required.
At a leaf $i$ with neighbor $i_1$ in $T$, the power assignment is $p_{ii_1}$ if $i$ also is the message source, and zero otherwise.

All nodes in $D$ are assumed to be equally likely to initiate a message.
The frequency at which power levels $p_{ii_1}$ and $p_{ii_2}$ (let $p_{ii_2}=0$ if $i$ is a leaf) are required,
is thus proportional to the number of sources in $D$ that utilize the respective arc for message forwarding.
In the example in Fig.\ \ref{fig:objexp}, the expected power consumption at node $i$ is thus proportional to $4p_{ii_1} + 3p_{ii_2}$.
Summing over all nodes yields the total cost of the tree.
% The objective in the problem addressed in the current work is to minimize this cost.

To express the cost of $T$ more rigorously, let $T^s=(V_T,A^s)$ denote the directed tree (arborescence) obtained by directing all edges in $E_T$ such that they point away from the \emph{root} at $s$.
For each arc $(i,j)\in A$, we let $n_{ij}(T)$ denote the number of destinations $s$ such that $(i,j)$ is the most expensive arc leaving node $i$ in $T^s$.
That is,
\[
  n_{ij}(T) = \left|\left\{s\in D: p_{ij}=\max_{k\in V_i: (i,k)\in A^s}p_{ik}\right\}\right|,
\]

\noindent
and the total cost of tree $T$ becomes:
$$
  c(T) = \sum_{i\in V_T}\sum_{j\in V_T:\{i,j\}\in E_T}p_{ij}n_{ij}(T).
$$ 
The problem under study is thus defined in network optimization terms:

\begin{problem}
\label{def:problem}
\textsc{Minimum Shared Multicast Tree}: Find a tree $T\subseteq G$ spanning $D$ such that $c(T)$ is minimized.
\end{problem}

While the feasible domains of the \textsc{Minimum Steiner Tree} problem and SMT are identical,
the cost functions differ.
In the former problem, all edges present in the tree contribute to the objective function by a cost that is given by the input data.
The input cost factor $p_{ij}$ is, in the case of SMT, scaled by a topology dependent factor $n_{ij}(T)$.
Only edges that are the most or second-most expensive edge incident to either of their end nodes get a positive scaling factor.
Some edges in $T$ can thus be scaled down to zero cost, while the most expensive ones are scaled up.

Problem \ref{def:problem} also has close resemblance with the SBT problem, which is the special case of SMT where $D=V$.
Like many other network design problems, \textsc{Minimum Shared Multicast Tree} is NP-hard.
This follows directly from the NP-hardness of SBT \citep{Papadimitriou06SBT}.

In the forthcoming sections, we develop and analyze IP models for SMT.
All formulations under study are extensions of models previously developed for the related problems mentioned above.