### 2018-12-28 smt/response2.tex
Response letter to the 2nd revision.

\documentclass[12pt]{article}
\usepackage{a4wide,url}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{color}

\begin{document}
\section*{Response Letter}
We are grateful to the editor and the reviewers for their time, effort, and constructive remarks in the review process.
We are also glad that the reviewers are satisfied with our first revision, and that they recommend acceptance conditioned upon correction of typos.

The manuscript has been revised for a second time, according to the reviewers' suggestions:
We have proofread it carefully, with particular attention to English grammar and spelling, and made a few corrections.
We hope that the editor finds the revision version suitable for publication in \emph{Journal of Combinatorial Optimization}.

\end{document}